---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

Set working directory for my computer
```{r}
setwd("C:/Users/canyon/Documents/GitHub/DataScience/Reproducible Research/RepData_PeerAssessment1-master")
```

Load csv
```{r}
steps = read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Load packages
```{r}
library(ggplot2)
library(lubridate)
```

Split date column into day, month, year
```{r}
steps$day = day(steps$date)
steps$month = month(steps$date)
steps$year  = year(steps$date)
```

The mean of the totals steps taken per day is below
```{r}
mean(steps$steps, na.rm = T)
median(steps$steps, na.rm = T)
```

## What is the average daily activity pattern?
```{r}
library(plyr)
avgsteps = ddply(steps, .(interval), summarise, msteps = mean(steps, na.rm = T))
qplot(interval, msteps, data = avgsteps)
```

## Imputing missing values
All missing values will equal global average
```{r}
steps$steps[is.na(steps$steps)] = mean(steps$steps, na.rm = T)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}

weekend <- function(day){
  if (wday(day, label = T) %in% c("Sun", "Sat")){
    return("Weekend")
    }
  return("Weekday")
}
steps = ddply(steps, .(date,interval), transform, day = weekend(date))

qplot(interval, steps, data = steps, facets = day~.)
```


