---
title: "Assignment 2"
author: "Ravin"
date: "Friday, September 05, 2014"
output: html_document
---

This analysis used storm data available onlien and condensed it to provide concise recommendations as to where to allocate resources. Summary analysis were performed to determine the most deadly and costly weather events across the United States. It was found that tornadoes by far were the deadliest weather event

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r DownloadFiles, eval = FALSE}
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile = "storm.bz2")
```

```{r LoadData, echo=FALSE, eval = FALSE}
storm = read.csv(bzfile("storm.bz2"))
```

```{r DamageFunction, echo = FALSE}
fdamage = function(x){
  #Set Multipliers
  factorlevel = c("","-","?","+","0","1","2","3","4","5","6","7","8","B","h","H","K","m","M")
  
  #Some of the factor levels were not known and were assumed to be zero
  factormultiplier = c(0,0,0,0,0,1,2,3,4,5,6,7,10^9,100,100,1000,1000,10^6,10^6)
  
  names(factormultiplier) = factorlevel
  return(factormultiplier[[x]])
}
```

```{r CostofDamage, echo = FALSE}

damage <- (storm[["CROPDMG"]]*sapply(storm[["CROPDMGEXP"]],fdamage) 
          +storm[["PROPDMG"]]*sapply(storm[["PROPDMGEXP"]],fdamage))

storm = cbind(storm,damage)
rm(damage)
```

Data by event type and fatality
```{r FatalitiesSummary, echo = FALSE}
library(plyr)
ef = ddply(storm, .(EVTYPE), summarise, fatalities = sum(FATALITIES), injuries = sum(INJURIES), affected = (sum(FATALITIES)+sum(INJURIES)), damage = sum(damage))

ef = ef[order(ef$affected, decreasing = TRUE),]
```


```{r Diasters}
library(reshape2)
library(ggplot2)
dfm = melt(data = ef[1:10,], measure.vars = c("fatalities","injuries","affected"))

ggplot(dfm, aes(reorder(EVTYPE,-value), value, fill = variable)) + geom_bar(position = "dodge", stat = "identity") + scale_x_discrete(labels = abbreviate) + ylab("Number of people") + xlab("Diasters") + ggtitle("People affected by diaster")
```


```{r Costs}
ef = ef[order(ef$damage, decreasing = TRUE),]
ggplot(ef[1:10,], aes(x = reorder(EVTYPE,-damage),damage))+ geom_bar(position = "dodge", stat = "identity") + scale_x_discrete(labels = abbreviate) + ylab("Cost of Diaster") + xlab("Type of Diaster") + ggtitle("Cost of diaster per Event Type")
```



#Appendix
Download the Storm data
```{r DownloadFiles, eval = FALSE}
```
Load storm data into dataframe
```{r LoadData, eval = FALSE}
```
Create a function with estimates damage
```{r DamageFunction, eval = FALSE}
```

Add cost of damage column to dataframe
```{r CostofDamage, eval = FALSE}
```



